Staged Abstract Interpreters

### 2019-09-11 paper/60_cases.tex
typo
polish

We show AC can be understood and implemented as an instance of staging abstract
interpreters. We first revisit the original implementation of abstract
compilation for 0-CFA, and then reproduce their result by simply adding stage
annotations.  The program generated by our approach provides approximately the
same improvement with regards to speed, but without changing a single line of
the analyzer program using type-based staging annotations. However, AC requires
more engineering effort, i.e., rewriting the whole analyzer into the
closure-generation form. Moreover, as shown in Section~\ref{staged_ds}, our
approach is able to not only remove the interpretive overhead, but also enable
several optimizations such as specialized data structures.

@@

Every time when we call the @eval@ function, we also refresh the timestamp by
calling the @tick@ function, which updates the timestamp in the state monad.
The @tick@ function is implemented as prepending the current control expression
being evaluated (@e@) to the existing calling context, and then taking the
first $k$ elements from the list. The type @Config@ used in the caching
algorithm is also changed to include timestamps.

>>>>

We show that AC can be understood and implemented as an instance of staging abstract
interpreters. We first revisit the original implementation of abstract
compilation for 0-CFA, and then reproduce their result by simply adding stage
annotations.  The program generated by our approach provides approximately the
same improvement with regards to speed, but without changing a single line of
the analyzer program using type-based staging annotations. However, AC requires
more engineering effort, i.e., rewriting the whole analyzer into the
closure-generation form. Moreover, as shown in Section~\ref{staged_ds}, our
approach is able to not only remove the interpretive overhead, but also enable
several optimizations such as specialized data structures.

@@

Every time when we call the @eval@ function, we also refresh the timestamp by
calling the @tick@ function, which updates the timestamp in the state monad.
The @tick@ function is implemented as prepending the current control expression
@e@ being evaluated to the existing calling context, and then taking the first
$k$ elements from the list. The type @Config@ used in the caching algorithm is
also changed to include timestamps.

### 2019-09-11 paper/80_related.tex
polish

Splitting an analysis into multiple stages has also been studied for other
forms of static analyses.
\citet{DBLP:conf/cgo/HardekopfL11} applied staging to a
flow-sensitive pointer analysis. The first stage is to analyze the
programs and obtain a sparse representation, and then the second
stage conducts a flow-sensitive analysis by refining the first
one. Many other forms of static analysis are implemented by
first generating constraints and then using a solver to discharge the
proof obligations or find the fixed-point. Notable examples include
SAT- and SMT-based program analysis
\cite{Gulwani:2008:PAC:1375581.1375616} and Datalog-based points-to
analysis \cite{Smaragdakis:2015:PA:2802194.2802195}.
Such analyses can be considered having stage distinctions.  One piece of future
work is to explore whether they can be implemented and optimized using
meta-programming.
Souffl{\'e}, one of the state-of-the-art Datalog engines commonly used in
points-to analysis \cite{Antoniadis:2017:PDS:3088515.3088522}, uses Futamura
projections for efficient implementation \cite{10.1007/978-3-319-41540-6_23}.
However, it only specializes the solving engine to Datalog rules, instead of a
full pipeline from a program AST to a specialized solver.

>>>>

Splitting an analysis into multiple stages has also been studied for other
forms of static analyses.
\citet{DBLP:conf/cgo/HardekopfL11} applied staging to a
flow-sensitive pointer analysis. The first stage is to analyze the
programs and obtain a sparse representation, and then the second
stage conducts a flow-sensitive analysis by refining the first
one. Many other forms of static analysis are implemented by
first collecting constraints from programs and then using a solver to discharge
the proof obligations or find the fixed-point. Notable examples include SAT-
and SMT-based program analysis
\cite{Gulwani:2008:PAC:1375581.1375616} and Datalog-based points-to
analysis \cite{Smaragdakis:2015:PA:2802194.2802195}.
Such analyses can be considered having stage distinctions.  One piece of future
work is to explore whether they can be implemented and optimized using
meta-programming.
Souffl{\'e}, one of the state-of-the-art Datalog engines commonly used in
points-to analysis \cite{Antoniadis:2017:PDS:3088515.3088522}, uses Futamura
projections for efficient implementation \cite{10.1007/978-3-319-41540-6_23}.
However, it only specializes the solving engine to Datalog rules, instead of a
full pipeline from a program AST to a specialized solver.

### 2019-09-11 paper/paper.tex
polish

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  We thank the reviewers of PLDI '19 and OOPSLA '19, as well as James Decker,
  Guanhong Tao, Gr\'egory Essertel, and Fei Wang for their feedback on the early
  drafts. We thank David Van Horn and David Darais for their encouragement on
  this work.
  This work was supported in part by NSF awards 1553471, 1564207, 1918483, DOE
  award DE-SC0018050, as well as gifts from Google, Facebook, and VMware.

>>>>

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  We thank the reviewers of PLDI '19 and OOPSLA '19, as well as James Decker,
  Guanhong Tao, Gr\'egory Essertel, Fei Wang, Qianchuan Ye, Yapeng Ye, and Le Yu
  for their feedback on the early drafts. We thank David Van Horn and David
  Darais for their encouragement on this work.
  This work was supported in part by NSF awards 1553471, 1564207, 1918483, DOE
  award DE-SC0018050, as well as gifts from Google, Facebook, and VMware.
